---
title: "Initial Korean covid 19"
author: "Angelica Lopez"
date: "Summer 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  

# Your Question - Is dying independent of sex?
##Method   
today we will analyze the Sex of the patient and the result whether or not theyre alive or dead. i will use the variables in the collection of Cat~cat. this is a type of problem that we can use our helper file for, to copy into our project and get the correct results.  
  



## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```  
this graph is showing me that in the one on the far left is saying the most people survived and that there were only a small number of people dead when going by the count of people. the second graph on the far right is showing us the percentages of these people but this graph is showing that far more people died, about 37% of women died and about 63% of men died, majoritey of deaths we made by men.    
    





  
## Numerical results


## Inferential results 




## Conclusions

