---
title: "Initial Korean covid 19"
author: "Angelica Lopez"
date: "Summer 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  

# Your Question - Is dying independent of sex?
##Method   
today we will analyze the Sex of the patient and the result whether or not theyre alive or dead. i will use the variables in the collection of Cat~cat. this is a type of problem that we can use our helper file for, to copy into our project and get the correct results.  
  



## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```  

this graph is showing me that in the first one, most people survived and that there were only a small number of people dead when going by the count of people. the second graph is showing us the percentages of these people but this graph is showing that far more people died, about 37% of women died and about 63% of men died, majority of deaths we made by men.    
    
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```  

the first chart says that of all the people that are still alive 56% are female and 44% are males. in the people that are dead 36% are females and 64% are men. this shows that the dead category is dominating with the deaths of males.another way to look at this, as stated in the secpnd table, 98.7% of women of still alive while 1.4% are dead and 97% of males are alive and 3% are dead. although the deaths arent a big percent, the male deaths are a little over more than half of the female deaths     

## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```  

in the group of people who are still alive there are 1831 females and 1411 males, and dead there are 25 females and 44 males. the p-value is 12/1000 the chi square test says . According to the Fishers Exact Test for count Data, the odds of females living is 2.3 times greater than the males living. males are more likely to die of covid than females are     
  
## Conclusions

in todays class we used our cat~cat helper file to assist us while trying to solve this problem. we started off by forking a code and putting that code into our posit and starting the assignment by putting our credentials and changing the name and date and committing. we then went on to run the first code given to us to open up our data in the environmental tab. Next, we created a method for what we were about to do and how we were going to do it. we followed that by opening out helper file cat~cat to retrieve our first code for our graphs, we got the code, changed it to match our data and ran that for our graphs, we reviewed and explained what we saw and comitted. we went back to our helper file to retrieve a code for our numerical results to give us our table with our data put into numerical form. we reviewed, explained, and comitted. finally, for our inferential results we used our helper file for the last time to get a code for our table. we ran that code and got our results. we came to the conclusion that the deaths of males due to covid are higher than those of women in Korea.   